---
title: "Models for Analyses"
author: "Nikolina Klatt"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{=html}
<style>
  div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
      </style>
```

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

```{r}
library(tidyverse) 
library(readxl) 
library(jtools)
library(janitor)
library(ggplot2)
library(broom)
library(gt)
library(plotrix)
library(ggeffects)
library(sjPlot)
library(gtsummary)
library(mgcv)
library(survival)
library(fixest)
library(lme4)
library(plm)
library(effects)
library(ggeffects)
library(webshot2)
library(Hmisc)
```

## Data Preparation

```{r}
tweets_final <- read_excel("data/tweets_final.xlsx")
    
# select relevant variables 
data <- tweets_final %>% 
  dplyr::select(date, docnum, text, name, state, party, congress, pred_story_type) %>% 
  filter(party != "I")
    
# create new relevant variables 
data <- data %>%
  mutate(overturn = as.factor(ifelse(date >= "2022-06-24", 1, 0)),
         story_type = as.factor(pred_story_type),
         story_type_binary = ifelse(pred_story_type == "Story of Decline", 1, 0),
         party = as.factor(party),
         party_binary = if_else(party == "R", 1, 0),
         state_trigger_law = as.factor(as.integer(state %in% c(
          "AL", "AZ", "AR", "ID", "KY", "LA", "MS", "MO", 
          "ND", "OK", "SD", "TN", "TX", "UT", "WY", "WV", "WI"))),
         date = as.Date(date),
         date_numeric = as.numeric(date),  
         date_for_plot = as.Date(date_numeric, origin = "1970-01-01"),
         events = factor(case_when(
           date < "2022-05-02" ~ "Before Leak",
           date >= "2022-05-02" & date < "2022-06-24" ~ "Between Leak and Overturning",
           date >= "2022-06-24" & date < "2022-09-13" ~ "Between Overturning and Bill",
           date >= "2022-09-13" & date < "2022-11-08" ~ "Between Bill and Midterm Elections",
           date >= "2022-11-08" ~ "After the Midterm Elections"), 
           levels = c("Before Leak", 
                      "Between Leak and Overturning", 
                      "Between Overturning and Bill",
                      "Between Bill and Midterm Elections",
                      "After the Midterm Elections"
                      ))
         )
```

### Figure 1: Frequency Plot

```{r}
fig_1 <- ggplot(data %>% filter(!party=="I"), aes(x = date, fill = party)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_vline(xintercept = as.Date("2022-05-02"), color = "grey", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-06-24"), color = "grey", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-09-13"), color = "grey", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-11-08"), color = "grey", linetype = "dashed") +
  annotate("text", x = as.Date("2022-05-02"), y = 160, label = "Leaked Draft", color = "grey", angle = 90, vjust = -0.5) + 
  annotate("text", x = as.Date("2022-06-24"), y = 163, label = "Overturning", color = "grey", angle = 90, vjust = -0.5) +
  annotate("text", x = as.Date("2022-09-13"), y = 145, label = "15-Week Ban Proposal", color = "grey", angle = 90, vjust = -0.5) +
  annotate("text", x = as.Date("2022-11-08"), y = 155, label = "Midterm Elections", color = "grey", angle = 90, vjust = -0.5) +
  labs(title = "Frequency Distribution of Abortion-related Tweets by Party Affiliation",
       x = "",
       y = "Number of Tweets",
       subtitle = "",
       fill = "Party") +
  scale_fill_manual(values = c("#0072B2", "red")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("output/fig_1.png", plot = fig_1, width = 10, height = 6, dpi = 300)

fig_1
```

## H0: No effect of the overturning of Roe v. Wade on the story types used in narratives by US representatives on Twitter

### Table 2: Models 1 and 2

```{r}
# Create Model 1
model_1 <- glm(story_type_binary ~ overturn,
               family = binomial(),
               data = data)

exp(coef(model_1))

# Create regression table for model 1
table_model_1 <- model_1 %>%
  tbl_regression(exponentiate = FALSE) %>%
  bold_labels() %>%
  bold_p(t = .1) %>%
  add_glance_table(include = nobs) 
    
# Create Model 2
model_2 <- glmer(story_type_binary ~ events +
                   (1 | name),
                 data = data,
                 family = binomial())

# Create regression table for model 2
table_model_2 <- model_2 %>%
  tbl_regression(exponentiate = FALSE) %>%
  bold_labels() %>%
  bold_p(t = .1) %>%
  add_glance_table(include = nobs)
    
# Merge the tables for models 1 and 2
table_merged_1_2 <- tbl_merge(
  list(table_model_1, table_model_2),
  tab_spanner = c("**Logistic Regression Model (1)**", 
                  "**Generalized Linear Mixed Model (2)**")
    )
    
# Convert to gt for additional customization
table_1_2_gt <- table_merged_1_2 %>%
      as_gt() %>%
      tab_spanner(
        label = "",
        columns = starts_with("estimate")
      ) 

gtsave(table_1_2_gt, filename = "output/table_1.png")

table_1_2_gt
```

### Figure 2: Marginal Effects Model 2

```{r}
# Effects of Model 2
model_2_effects <- effect("events", model_2)
    
# Convert effects object to a data frame
effects_df <- as.data.frame(model_2_effects)
    
# Plot marginal effects 
fig_2 <- ggplot(effects_df, aes(x = events, y = fit)) +
  geom_point() +  # Add points to see data points
  geom_line(group = 1) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, alpha = 0.5, color = "red") +  # This adds the error bars
  labs(title = "Marginal Effects of Story of Decline",
       subtitle = "Generalized Linear Mixed Model (2)",
       x = "",
       y = "Estimated Effect") +
  theme_minimal()
    
ggsave("output/fig_2.png", fig_2, width = 11, height = 5, dpi = 300)

fig_2
```

### Figure 3: Generalized Additive Model 3 - Time Aspect

```{r}
# GAM model with date_numeric
model_3 <- gam(story_type_binary ~ s(date_numeric) +
                 events +
                 party +
                 state_trigger_law,
               family = binomial(),
               data = data)
    
# Create plot data from the GAM model
model_3_plot_data <- data.frame(
  date_for_plot = data$date_for_plot,
  party = data$party,
  fitted_values = fitted(model_3),
  se_fit = sqrt(summary(model_3)$s.table[, "edf"]) * 2)
    
# Plot 
fig_3 <- ggplot(model_3_plot_data, aes(x = date_for_plot, y = fitted_values)) +
  geom_line() +
  geom_vline(xintercept = as.Date("2022-05-02"), color = "red", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-06-24"), color = "red", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-09-13"), color = "red", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2022-11-08"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2022-05-02"), y = 0.55, label = "Leaked Draft", color = "red", angle = 90, vjust = -0.5) +
  annotate("text", x = as.Date("2022-06-24"), y = 0.55, label = "Overturning", color = "red", angle = 90, vjust = -0.5) +
  annotate("text", x = as.Date("2022-09-13"), y = 0.45, label = "15-Week Ban Proposal", color = "red", angle = 90, vjust = -0.5) +
  annotate("text", x = as.Date("2022-11-08"), y = 0.57, label = "Midterm Elections", color = "red", angle = 90, vjust = -0.5) +
  labs(title = "Change in Probability of Story of Decline",
       subtitle = "Generalized Additive Model (3)",
       x = "", 
       y = "Fitted Values") +
  theme_minimal()
    
ggsave("output/fig_3.png", fig_3, width = 9, height = 5, dpi = 300)

fig_3
```

## H1: Democrats and Republicans exhibited different adjustments in their story types after the overturning.

### Table 2: Models 4 and 5

```{r}
# Create Model 4
model_4 <- glmer(story_type_binary ~ party*events +
                       (1 | name),
                     data = data,
                     family = binomial())

# Create regression table for model 4
table_model_4 <- model_4 %>%
  tbl_regression(exponentiate = FALSE) %>%
  bold_labels() %>%
  bold_p(t = .1) %>%
  add_glance_table(include = nobs)

# Create dataset for Republicans for Model 5
rep <- data %>%
  filter(party == "R")

# Create Model 5
model_5 <- glmer(story_type_binary ~ events +
                   (1 | name),
                 data = rep,
                 family = binomial())
    
# Create regression table for model 5
table_model_5 <- model_5 %>%
  tbl_regression(exponentiate = FALSE) %>%
  bold_labels() %>%
  bold_p(t = .1) %>%
  add_glance_table(include = nobs)
    
# Merge the tables
table_merged_4_5 <- tbl_merge(
  list(table_model_4, table_model_5),
  tab_spanner = c("**GLMM (4)**", 
                  "**GLMM Republicans (5)**")
    )
    
# Convert to gt for additional customization
table_4_5_gt <- table_merged_4_5 %>%
  as_gt() %>%
  tab_spanner(label = "",
              columns = starts_with("estimate")
      ) 
    
gtsave(table_4_5_gt, filename = "output/table_2.png")

table_4_5_gt
```

### Figure 4: Predicted Probability Model 4

```{r}
# Calculating predicted probabilities
model_4_pp <- ggpredict(model_4, terms = c("events", "party"))
    
fig_4 <- ggplot(model_4_pp, aes(x = x, 
                                y = predicted, 
                                color = group, 
                                group = group)) +
  geom_line() +
  geom_point() +
  labs(title = "Predicted Probability of 'Story of Decline' by Party",
       subtitle = "Generalized Linear Mixed Model (4)",
       x = "",
       y = "Predicted Probability",
       color = "Party") +
  scale_color_manual(values = c("#56B4E9", "red")) +
  theme_minimal() +
  theme(legend.position = "bottom") 
    
ggsave("output/fig_4.png", fig_4, width = 11, height = 5, dpi = 300)

fig_4
```

## H2: Geographic variability in narrative shifts based on state-level abortion policies.

### Figure 5: Density of Proportion of Story of Decline

```{r}
state_data <- data %>%
      group_by(state, events, state_trigger_law) %>%
      summarise(proportion_decline = mean(story_type_binary, na.rm = TRUE)) %>% 
      ungroup()
    
custom_colors <- c("0" = "purple", "1" = "yellow")  
    
fig_5 <- ggplot(state_data, aes(x = proportion_decline, 
                                fill = as.factor(state_trigger_law))) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ events) +  
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Density of Proportion of Story of Decline by Trigger Law Status",
       subtitle = "Facetted by Events",
       x = "Proportion of Story of Decline",
       y = "Density",
       fill = "Trigger Law Status"  
      ) +
  scale_fill_manual(values = custom_colors)  
    
ggsave("output/fig_5.png", fig_5, width = 10, height = 6, dpi = 300)

fig_5
```

### Table 3: Models 6 and 7

```{r}
# Create Models 6 and 7
model_6 <- glmer(story_type_binary ~ state_trigger_law +
                   events +
                   party +
                   (1 | name),
                 data = data,
                 family = binomial())

model_7 <- glmer(story_type_binary ~ state_trigger_law*events +
                   state_trigger_law*party +
                   (1 | name),
                 data = data,
                 family = binomial())

# Create regression table for each model
table_model_6 <- tbl_regression(model_6, exponentiate = F) %>%
  bold_labels() %>% 
  bold_p(t = .1) %>%
  add_glance_table(include = nobs)
    
table_model_7 <- tbl_regression(model_7, exponentiate = F) %>%
  bold_labels() %>% 
  bold_p(t = .1) %>%
  add_glance_table(include = nobs)

# Merge the tables
table_merged_6_7 <- tbl_merge(
  list(table_model_6, table_model_7),
  tab_spanner = c("**GLMM (6)**", 
                  "**GLMM with Interactions (7)**")
    )
    
# Convert to gt for additional customization
table_6_7 <- table_merged_6_7 %>%
  as_gt() %>%
  tab_spanner(
    label = "",
    columns = starts_with("estimate")
      ) 

gtsave(table_6_7, filename = "output/table_3.png")

table_6_7
```
